---
title: "RSS-NET analysis of simulated GWAS summary statistics and B cell regulatory network"
author: "Xiang Zhu"
date: "2019-08-18"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

[`wtccc_bcell/`]: https://github.com/SUwonglab/rss-net/tree/master/examples/wtccc_bcell
[zenodo-wtcccbcell]: 

## Overview

Here we describe an end-to-end RSS-NET analysis of two synthetic datasets
that are simulated from real genotypes of 348965 genome-wide SNPs from
1458 individuals in the UK Blood Service Control Group
([Wellcome Trust Case Control Consortium, 2007](https://doi.org/10.1038/nature05911))
and a gene regulatory network inferred for B cells
[(ENCODE Project Consortium, 2012)](https://doi.org/10.1038/nature11247).
This example gives a quick view of how RSS-NET works.

To reproduce results of this example,
please use scripts in the directory [`wtccc_bcell/`][],
and follow the step-by-step guide below.
Before running any script in [`wtccc_bcell/`][],
please [install](setup.html) RSS-NET.

With the software installed and the input data downloaded,
one should be able to run this example by simply typing the following line in shell:

```{r, eval=FALSE, engine='zsh'}
$ sbatch run_simulation.sbatch
```

If a different directory is used to store the software, input data
and/or output data, please modify file paths in the given scripts accordingly.

## Step-by-step illustration

### Download data files

All data files required to run this example are freely available at Zenodo
[![DOI]()][zenodo-wtcccbcell].
Please contact me (`xiangzhu[at]stanford.edu`)
if you have any trouble accessing these files.
After a complete download, you should see the following files.

```{r, eval=FALSE, engine='zsh'}
add tree
```

To help readers confirm if they use the same files as we do,
we report 128-bit MD5 hashes of all files in `wtccc_bcell_data.md5`.

To help readers confirm if they can reproduce results of this example,
we also provide the results (`wtccc_bcell_results.zip`) in the same Zenodo deposit.

For simplicity and generality,
we introduce the following short-hand notations.

```matlab
gwas = 'WTCCC';
net  = 'Primary_B_cells_from_peripheral_blood'; 
```
